---
title: Underrated CRAN Packages
author: Mark Edney
date: '2022-03-31'
slug: []
categories:
  - Project
  - R
draft: false
image: 'Rlogo.svg'
description: 'A search for popular R packages, that I would would otherwise miss.'
archives:
  - 2022/03
---

![](Rlogo.svg)

I sit here looking for inspiration, nothing interesting to write about. Perhaps there are some popular R packages on CRAN that I don't know about? You can explore the data on downloads from CRAN with the `cranlogs` package. 

## Top CRAN downloads

With the following code, we can get the most popular packages from CRAN. The CRAN directory doesn't represent all R packages, but a good amount of them. 

```{r cran, warning=FALSE, message=FALSE}
library(tidyverse)
library(cranlogs)
top100 <- cran_top_downloads(when = 'last-month', count = 100)
top100 %>% head()
```
From this list, we can see that the `tidyverse` represents a large amount of the top downloads with `ggplot2`, `rlang` and `dplyr`. The list includes the `sf` package for geospacial data, the `glue` package for string manipulation and the `cli` package which is used to create a command line interface for packages. Most of these packages I already have a good understanding of, so I need to narrow down the search.

## Packages installed

You can get a list of your installed packages with the installed_packages function. You can then filter the top 100 list and remove anything you already have installed to find new packages. 

```{r packages}
mine <- installed.packages() %>%
        data.frame() %>%
        select(Package)
new <- top100 %>%
        filter(!package %in% mine$Package)
new
```

From some quick research, I have found the following about the new packages:

*  `ragg` - a 2D library as an alternative to the RStudio default
*  `rgl` - functions for 3D interactive graphics
*  `rgeos` - a geometry package, but is currently planned to be retired at the end of 2023 for the `sf` package 
*  `zoo` - a library to deal with time series
*  `pkgdown` - a library fOR building a blog website, I use blogdown
*  `nloptr` - a library for solving non-linear optimization problems
*  `Hmisc` - an assortment of different data analysis tools
*  `lme4` - for fitting linear and generalized linear mixed-effects models
*  `RcppEigen` - integration of the `eigen` library in R for linear algebra 

## Take-away

Hopefully your take-way is a simple method to explore R library that you have never heard about. I know that a few of the libraries seem interesting and worth further exploring. 

While we are at it, might as well find the daily values for the new packages and plot them for the last month. 

```{r plot}
new$package %>%
        cran_downloads(when = "last-month") %>%
        ggplot(aes(x = date, y = count, color = package)) +
        geom_line()
        
```